Draft Exploration

### 2020-04-03 draft_exploration.Rmd
made graphs of redraft probability and mean round change

unsigned_ids <- master %>%
  filter(source == "H" & signed == "N") %>%
  select(player_id)

@@

round_redraft_percent(2005, 2009, 5)

@@

redraft_table_50 <- tibble(start_year = c(1990, 1995, 2000, 2005, 2010)) %>%
  mutate(round = 50) %>%
  mutate(percent_redrafted = map_dbl(half_decade, ~round_redraft_percent(
    startyear = .,
    max_round = 50,
    min_round = 21), 
    .id = half_decade)) %>%
  mutate(initial_draft_round = map_dbl(half_decade, ~prior_mean_round(
    startyear = .,
    max_round = 50,
    min_round = 21),
    .id = half_decade)) %>%
  mutate(redraft_round = map_dbl(half_decade, ~post_mean_round(
    startyear = .,
    max_round = 50,
    min_round = 21),
    .id = half_decade)) %>%
  mutate(mean_round_diff = initial_draft_round - redraft_round)

redraft_table <- rbind(redraft_table_50, redraft_table_20, redraft_table_10, redraft_table_05)

@@

>>>>

# NOTE: I added the draft_numb == 1 to the filter because some players (Matt
# Harrington) were drafted out of high school technically and turned down the
# draft multiple times.

unsigned_ids <- master %>%
  filter(source == "H" & signed == "N" & draft_numb == 1) %>%
  select(player_id)

@@

round_redraft_percent(2000, 2004, 5)

@@

# I added the ifelse for rounds 40-50 for the average round columns because it
# is possible a player was redrafted after the new Collective Bargaining
# Agreement that cut the draft to 40 rounds. This would make it appear as though
# their jump in draft stock was higher than in reality because there were only
# 40 rounds to be selected from to begin with, implying an automatic draft stock
# boost if redrafted. I only used an ifelse to set 2010 data to NA for percent
# redrafted because we only have one year of data from 2011 in this category.

redraft_table_50 <- tibble(start_year = c(1990, 1995, 2000, 2005, 2010)) %>%
  mutate(round = 50) %>%
  mutate(percent_redrafted = ifelse(start_year == 2010, NA,
                                    map_dbl(half_decade, ~round_redraft_percent(
    startyear = .,
    max_round = 50,
    min_round = 41), 
    .id = half_decade))) %>%
  mutate(initial_draft_round = ifelse(start_year == 2010 | start_year == 2005, NA,
                                      map_dbl(half_decade, ~prior_mean_round(
    startyear = .,
    max_round = 50,
    min_round = 41),
    .id = half_decade))) %>%
  mutate(redraft_round = ifelse(start_year == 2010 | start_year == 2005, NA,
                                map_dbl(half_decade, ~post_mean_round(
    startyear = .,
    max_round = 50,
    min_round = 41),
    .id = half_decade))) %>%
  mutate(mean_round_diff = initial_draft_round - redraft_round)

redraft_table_30 <- tibble(start_year = c(1990, 1995, 2000, 2005, 2010)) %>%
  mutate(round = 30) %>%
  mutate(percent_redrafted = map_dbl(half_decade, ~round_redraft_percent(
    startyear = .,
    max_round = 30,
    min_round = 21), 
    .id = half_decade)) %>%
  mutate(initial_draft_round = map_dbl(half_decade, ~prior_mean_round(
    startyear = .,
    max_round = 30,
    min_round = 21),
    .id = half_decade)) %>%
  mutate(redraft_round = map_dbl(half_decade, ~post_mean_round(
    startyear = .,
    max_round = 30,
    min_round = 21),
    .id = half_decade)) %>%
  mutate(mean_round_diff = initial_draft_round - redraft_round)

redraft_table_40 <- tibble(start_year = c(1990, 1995, 2000, 2005, 2010)) %>%
  mutate(round = 40) %>%
  mutate(percent_redrafted = map_dbl(half_decade, ~round_redraft_percent(
    startyear = .,
    max_round = 40,
    min_round = 31), 
    .id = half_decade)) %>%
  mutate(initial_draft_round = map_dbl(half_decade, ~prior_mean_round(
    startyear = .,
    max_round = 40,
    min_round = 31),
    .id = half_decade)) %>%
  mutate(redraft_round = map_dbl(half_decade, ~post_mean_round(
    startyear = .,
    max_round = 40,
    min_round = 31),
    .id = half_decade)) %>%
  mutate(mean_round_diff = initial_draft_round - redraft_round)

redraft_table <- rbind(redraft_table_50,
                       redraft_table_40,
                       redraft_table_30,
                       redraft_table_20, 
                       redraft_table_10, 
                       redraft_table_05)

@@

# ggplot of redraft probability by set of rounds drafted in out of high school
# and 5 year period in which a player was drafted out of high school.

redraft_table %>%
  ggplot(aes(round, percent_redrafted)) +
  geom_point(na.rm = TRUE) +
  geom_line(aes(group = start_year, color = start_year), na.rm = TRUE) +
  labs(
    x = "Round Range Drafted out of High School",
    y = "Percent of Picks Re-drafted",
    color = "Time Frame",
    title = "High School Re-draft Probability",
    subtitle = "Organized by initial round range and time period"
  ) +
  scale_color_viridis_c() +
  scale_x_continuous(breaks = c(5, 10, 20, 30, 40, 50))
```

```{r}
redraft_table %>%
  ggplot(aes(round, mean_round_diff)) +
  geom_point(na.rm = TRUE) +
  geom_line(aes(group = start_year, color = start_year), na.rm = TRUE) +
  labs(
    x = "Round Range Drafted out of High School",
    y = "Average Improvement Re-draft Round",
    color = "Time Frame",
    title = "Average Change in Round of Re-drafted Unsigned High School Picks",
    subtitle = "Organized by initial round range and time period"
  ) +
  scale_color_viridis_c() +
  scale_x_continuous(breaks = c(5, 10, 20, 30, 40, 50)) +
  scale_y_continuous(breaks = c(-10, -5, 0, 10, 20)) +
  geom_hline(color = "red", linetype = "dashed", yintercept = 0)
```
